---
title: "HW8"
author: "Yining Qu"
date: "2024-05-15"
output: 
    pdf_document: default
---

```{r cars}
library(stats)
library(glmnet)
library(Rcpp) 
library(ggplot2) 
library(reshape2)
library(corrplot)
```

```{r pressure}
fx <- read.csv("FXmonthly.csv")
fx <- (fx[2:120,]-fx[1:119,])/(fx[1:119,])
sp500 <- read.csv("sp500.csv")
```

## Q1

```{r}
# Calculate and visualize the correlation matrix
cor_matrix <- cor(fx)
corrplot::corrplot(cor_matrix, method = "circle")
```
```{r}
print(cor_matrix)
```

## Q2
```{r}
# Fit PCA
pcfx <- prcomp(fx, scale = TRUE)

plot(pcfx, main="")

```
```{r}
pcs <- predict(pcfx)

plot(pcs[,1], pcs[,2], xlab = "First Principal Component", ylab = "Second Principal Component", main = "Scatter Plot of the First Two Principal Components", pch = 20, col = 'blue')

text(pcs[,1], pcs[,2], labels = rownames(pcs), pos = 4, cex = 0.6)

```
```{r}
plot(pcs[,2], pcs[,3], xlab = "Second Principal Component", ylab = "Third Principal Component", main = "Scatter Plot of the Second versus the Third Principal Components", pch = 20, col = 'blue')

text(pcs[,1], pcs[,2], labels = rownames(pcs), pos = 4, cex = 0.6)

```

```{r}
# far right
smallest_pcs <- sort(pcs[,1])[1:5]
smallest_pcs
```
```{r}
# far left
largest_pcs <- sort(pcs[,1], decreasing = TRUE)[1:5]
largest_pcs
```

```{r}
t(round(pcfx$rotation[,1:2],2))
```
PC1 Loadings:
- Mostly negative across many currencies, indicating a general pattern where an increase in the principal component score corresponds to a decrease in these currency values against their pair.
- The magnitude of the loadings (e.g., -0.28 for exdnus) signifies how much a unit change in the principal component affects the currency. Larger absolute values mean greater sensitivity to changes in this component.

PC2 Loadings:
- Mixed signs across different currencies indicate diverse influences captured by this component. For instance, a positive loading for exbzus (0.33) against a negative for exjpus (-0.46) highlights how different economic or market conditions might be driving movements in these currencies.

## Q3
```{r}
set.seed(4)

if (!requireNamespace("AICcmodavg", quietly = TRUE)) {
  install.packages("AICcmodavg")
}

# Load the AICcmodavg package
library(AICcmodavg)

# Convert the principal components to a data frame for modeling
zdf <- as.data.frame(pcs)

# Perform regression using the first two principal components
summary(fxglm <- glm(sp500$sp500 ~ ., data=zdf[, 1:2]))

# Function to calculate AICc for a glm model
calculateAICc <- function(model) {
  AICc(model)
}

# Perform GLM fits on 1:20 factors and calculate AICc for each
kfits <- lapply(1:20, function(K) {
  glm(sp500$sp500 ~ ., data = zdf[, 1:K, drop = FALSE])
})

aicc <- sapply(kfits, calculateAICc) # Apply AICc calculation to each fit
which.min(aicc) # Determine which model size (number of factors) is preferred by AICc

# Optionally, also calculate BIC for comparison
bic <- sapply(kfits, BIC)
which.min(bic) # Determine which model size (number of factors) is preferred by BIC
```
```{r}
# Plot the AICc values
plot(aicc, type = 'b', xlab = "Number of Factors", ylab = "AICc", main = "AICc by Number of Factors")
```

```{r}
best_aicc_index <- which.min(aicc)
print(coef(kfits[[best_aicc_index]]))
```



```{r}
# Ensure the glmnet package is installed and loaded
if (!requireNamespace("glmnet", quietly = TRUE)) {
    install.packages("glmnet")
}
library(glmnet)

# Assuming 'pcs' is a data frame or matrix of principal components
# and 'sp500$sp500' is a vector of your response variable

# Convert 'pcs' to matrix if it is not already
pcs_matrix <- as.matrix(pcs)

# Set seed for reproducibility

# Perform Lasso regression with cross-validation
lassoPCR <- cv.glmnet(x = pcs_matrix, y = sp500$sp500, family = "gaussian", nfolds = 20, alpha = 1)

# Coefficients at the best lambda (lambda that gives minimum mean cross-validated error)
lasso_coef <- coef(lassoPCR, s = "lambda.min")

# Plotting the results
par(mfrow=c(1,2))  # Set up the plotting area to have 1 row and 2 columns

# Plot AICc
plot(aicc, pch=21, bg="maroon", xlab="K", ylab="AICc", main="AICc by Number of Factors")

# Plot Lasso path
plot(lassoPCR)
title("Lasso Path")

```

```{r}
lambda_min <- lassoPCR$lambda.min
lasso_model_min <- glmnet(pcs_matrix, sp500$sp500, lambda = lambda_min)
coefficients_min <- coef(lassoPCR, s = "lambda.min")
print(coefficients_min)
```

## Q4
```{r}
# Prepare the data
fx_matrix <- as.matrix(fx)  # Convert fx to matrix if not already
sp500_vector <- sp500$sp500  # Assuming sp500$sp500 is already a vector

# Fit Lasso model using cross-validation to determine the best lambda
set.seed(41201)  # For reproducibility
lambda_values <- 10^seq(3, -2, by = -0.1)
cv_lasso_fx <- cv.glmnet(x = fx_matrix, y = sp500_vector, alpha = 1, family = "gaussian", nfolds = 20)

# Coefficients at the best lambda
best_lambda_fx <- cv_lasso_fx$lambda.min
coefficients_fx <- coef(cv_lasso_fx, s = "lambda.min")

# Print the coefficients
print(coefficients_fx)

# Plot the Lasso path
plot(cv_lasso_fx)
title("Lasso Path for Original Covariates")

```




